---
title: "DataVis Przybyłkiewicz"
output: 
  pdf_document
  
editor_options: 
  chunk_output_type: console
---
```{r, global_options, include=FALSE}
knitr::opts_chunk$set(message=FALSE) 
```

```{r, include=FALSE}
library(tidyverse)
library(tinytex)
library(RColorBrewer)
library(ggthemes)
library(extrafont)
font_import()

cces <- drop_na(read_csv(url("https://www.dropbox.com/s/ahmt12y39unicd2/cces_sample_coursera.csv?raw=1")))

cel <- drop_na(read_csv(url("https://www.dropbox.com/s/4ebgnkdhhxo5rac/cel_volden_wiseman%20_coursera.csv?raw=1")))
```

## Bar plot:

I wanted to show if there is disproportion in number of congress members based on their gender

```{r}
cel$party <- recode(cel$dem, `1`="Democrats", `0`="Republicans")
cel$gender <- recode(cel$female, `1`="Female", `0`="Male")
ggplot(cel, aes(x = party, fill = gender)) +
  geom_bar(position = "dodge2", width = 0.7) +
  labs (title = "Share of female and male congress members", x = NULL, y = "count", 
    fill = NULL) + scale_fill_manual(values = c("#E82CE4","#D9C1D0")) + 
  theme_minimal() + theme (text = element_text(family = "Nunito SemiBold"), plot.title = 
  element_text(face = "bold", size =12), panel.grid.major = element_line(colour = "black", 
  size = 0.1, linetype = "dotted"), panel.grid.minor = element_blank(), aspect.ratio = 1, 
  legend.text = element_text(size = 8), legend.key.size = unit(0.8, "lines"), 
  legend.box.spacing = unit(0.1, "cm"), axis.text.y = element_text(size = 7)) +
  scale_y_continuous(limits = c(0,6000))
```

## Scatter plot:

I've decided to stick with marginalized groups and visualize how many bills proposed by members with different levels of seniority became laws, grouping them by gender and ethnicity

```{r}
cel$etni <- recode(cel$afam, `0`="Other", `1`="African American")
cel$membership <- recode(cel$dem, `0`="Republican", `1`="Democrat")

cel %>% filter(year>=2010) %>% ggplot(aes(x=seniority,y=all_law,
                                          color=membership)) +
  geom_point(position = "jitter", alpha=1/2) +
  facet_wrap(gender~etni, labeller = label_wrap_gen(multi_line=FALSE)) +
  labs(x="seniority", y="bills transformed into laws", 
       color="Political Party") +
  scale_color_manual(values = c("#2C3AE8", "#E82CE4")) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold"),
    axis.title.x = element_text(size=10, hjust = 1),
    axis.title.y = element_text(size=10, hjust = 1),
    legend.position = "top",
    legend.title = element_blank(), 
    legend.justification = c(-0.08,0), 
    legend.margin = margin(0), 
    aspect.ratio = 1) + 
  ggtitle("relation of member seniority to bills passed ", 
          subtitle = "by ethnicity and gender") 
```

## Line plot:

In Exercise 1 I was checking if there is a disproportion and now I want to examine its division by ethnicity

```{r}
ex3dat <- cel %>% count(year, etni, gender)
annotation <- data.frame(ex3dat, lab = "disproportion between genders") %>% 
  filter(etni=="Other")

ggplot(ex3dat, aes(x=year, y=n, color=gender)) +
  geom_line() + 
  geom_text(data=annotation, aes(x=1995,y=200,label=lab), 
            color="black", check_overlap = TRUE) +
  facet_wrap(~etni) +
  labs(title="Gender division in congress by ethnicity", 
       x="Year of observation", y="Number of Indyviduals") + 
  scale_color_manual(values = c("#E82CE4","#2C3AE8")) +
  theme_bw() +
  theme(legend.title = element_blank())
```

## Density plot:

Visualization of ideology distribution among congress members divided by ethnicity and gender after year 2000

```{r}
caption <-paste(
  strwrap("Ideology distribution of congress members after year 2000", 40), 
  collapse = "\n")

cel %>% filter(year>=2000) %>% ggplot(aes(x=dwnom1, color=gender)) +
geom_density() + 
  annotate(geom = "text", x = 0, y = 3, label = caption, hjust = 0, size = 3) +
  facet_grid(etni~ .) +
  labs(x="ideology (higher = more conservative)", y="density") +
  scale_color_manual(values = c("#E82CE4","#2C3AE8")) +
  theme_minimal() +
  theme(panel.spacing = unit(1, "lines"),
    plot.title = element_text(face = "bold"),
    axis.title.x = element_text(size=10),
    axis.title.y = element_text(size=10),
    legend.position = "right",
    legend.title = element_blank(), legend.box = "dot") 
```
